---
title: "Mount_Analysis"
author: "Jay Falk"
date: "5/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages
```{r load-library, message = FALSE}
library(tidyverse)
```

Load stored datasets
```{r datasets, message = FALSE}
longcomplete <- read_csv("FirstReaction_gittrack.csv")
behaviors <- read_csv("AllBehaviors_gittrack.csv")
trialsides_sides <- read_csv("trialsides_gittrack.csv")
```

##First Reaction Interactions with Mounts

Data shaping for analysis
```{r}
#First filter out the trials where no behaviors happened at all
nona <- longcomplete %>%
  mutate(sumthing = if_else(feed == 0 & attack == 0 & tail == 0 & court == 0 & cop == 0, 0, 1)) %>%
  filter(sumthing == 1) %>%
  select(-sumthing)

#Make some categories based on the presence of certain types
testcategories_nona <- nona %>%
  mutate(Fpresent = if_else(test == "M/MLF", 0, 1)) %>%
  mutate(Mpresent = if_else(test == "F/MLF", 0, 1)) %>%
  mutate(MLFpresent = if_else(test == "M/F", 0, 1))
```

Let's first look at sexual behavior
```{r}
courtscops_nona <- testcategories_nona %>%
  mutate(sumthing = if_else(court == 0 & cop == 0, 0, 1)) %>%
  filter(sumthing == 1) %>%
  select(date, trial, test, type, court, cop)

replace_na(courtscops_nona$court,0)
replace_na(courtscops_nona$cop,0)
```

Feed behavior. First look at whether there are differences between types of interactors in who they choose to feed with. Types are androchromes, heterochromes, Amazilias, and Other
```{r}
#Do types of birds differ in the way they interact with males versus heterochrome females? 
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(feed != 0)
mffeedtable <- table(mf$type, mf$feed)
prop.table(mffeedtable, m=1)
chisq.test(mffeedtable)

#Do types of birds differ in the way they interact with androchrome females versus heterochrome females? 
fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(feed != 0)
fmlffeedtable <- table(fmlf$type, fmlf$feed)
prop.table(fmlffeedtable, m=1)
chisq.test(fmlffeedtable)

#Do types of birds differ in the way they interact with androchrome females versus males? 
mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(feed != 0)
mmlffeedtable <- table(mmlf$type, mmlf$feed)
prop.table(mmlffeedtable, m=1)
chisq.test(mmlffeedtable)
```

Are mounts fed with at different rates? Androchromes seem to be doing something different than the other types of birds in the male versus heterochrome female test type. See what happens if we take them out. 
```{r}
#Do all hummingbirds feed with male versus heterochrome female mounts at different rates? 
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(feed != 0)
mffeedtable <- table(mf$feed)
prop.table(mffeedtable)
chisq.test(mffeedtable)

#What about if we take out the androchromes?  
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(feed != 0) %>%
  filter(type != "Flom")
mffeedtable <- table(mf$feed)
prop.table(mffeedtable)
chisq.test(mffeedtable)

#What about if we take out jacobins?  
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(feed != 0) %>%
  filter(type != "Flom", type != "Flof")
mffeedtable <- table(mf$feed)
prop.table(mffeedtable)
chisq.test(mffeedtable)



#Do all hummingbirds feed with androchrome female versus heterochrome female mounts at different rates? 
fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(feed != 0)
fmlffeedtable <- table(fmlf$feed)
prop.table(fmlffeedtable)
chisq.test(fmlffeedtable)

#What about if we take out the androchromes?  
fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(feed != 0) %>%
  filter(type != "Flom")
fmlffeedtable <- table(fmlf$feed)
prop.table(fmlffeedtable)
chisq.test(fmlffeedtable)

#What about if we take out all jacobins?  
fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(feed != 0) %>%
  filter(type != "Flom", type != "Flof")
fmlffeedtable <- table(fmlf$feed)
prop.table(fmlffeedtable)
chisq.test(fmlffeedtable)



#Do all hummingbirds feed with males versus androchrome female mounts at different rates? 
mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(feed != 0)
mmlffeedtable <- table(mmlf$feed)
prop.table(mmlffeedtable)
chisq.test(mmlffeedtable)

#What about if we take out the androchromes?  
mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(feed != 0) %>%
  filter(type != "Flom")
mmlffeedtable <- table(mmlf$feed)
prop.table(mmlffeedtable)
chisq.test(mmlffeedtable)

#What about if we take out all jacobins?  
mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(feed != 0) %>%
  filter(type != "Flom", type != "Flof")
mmlffeedtable <- table(mmlf$feed)
prop.table(mmlffeedtable)
chisq.test(mmlffeedtable)
```


Ok, so that's what non-androchromes do. What about androchromes?
```{r}
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(feed != 0) %>%
  filter(type == "Flom")
mffeedtable <- table(mf$feed)
prop.table(mffeedtable)
chisq.test(mffeedtable)
#         F         M 
# 0.4137931 0.5862069 

fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(feed != 0) %>%
  filter(type == "Flom")
fmlffeedtable <- table(fmlf$feed)
prop.table(fmlffeedtable)
chisq.test(fmlffeedtable)
#         F       MLF 
# 0.4666667 0.5333333 

mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(feed != 0) %>%
  filter(type == "Flom")
mmlffeedtable <- table(mmlf$feed)
prop.table(mmlffeedtable)
chisq.test(mmlffeedtable)

#         M       MLF 
# 0.5555556 0.4444444 
```


Attack behavior. First look at whether there are differences between types of interactors in who they choose to attack. Types are androchromes, heterochromes, Amazilias, and Other.
```{r}
#Do types of birds differ in the way they attack males versus heterochrome females? 
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(attack != 0)
mfattacktable <- table(mf$type, mf$attack)
prop.table(mfattacktable, m=1)
fisher.test(mfattacktable)

#Do types of birds differ in the way they attack androchrome females versus heterochrome females? 
fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(attack != 0)
fmlfattacktable <- table(fmlf$type, fmlf$attack)
prop.table(fmlfattacktable, m=1)
fisher.test(fmlfattacktable)

#Do types of birds differ in the way they attack androchrome females versus males? 
mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(attack != 0)
mmlfattacktable <- table(mmlf$type, mmlf$attack)
prop.table(mmlfattacktable, m=1)
fisher.test(mmlfattacktable)

test <- testcategories_nona %>% filter(attack != 0) 
testtable <- table(test$type)

```

Compile all the attacks since no evidence that types are doing different things, then test whether mounts are treated differently
```{r}
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(attack != 0)
mfattacktable <- table(mf$attack)
prop.table(mfattacktable)
chisq.test(mfattacktable)
binom.test(mfattacktable)

# 	Chi-squared test for given probabilities
# 
# data:  mfattacktable
# X-squared = 6.8182, df = 1, p-value = 0.009023

mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(attack != 0) %>%
  filter(type != "Flom")
mfattacktable <- table(mf$attack)
prop.table(mfattacktable)
chisq.test(mfattacktable, correct = F)
binom.test(mfattacktable, p = 0.5)

# 	Chi-squared test for given probabilities
# 
# data:  mfattacktable
# X-squared = 4.5714, df = 1, p-value = 0.03251

mf <- testcategories_nona %>%
  filter(test == "M/F") %>%
  filter(attack != 0) %>%
  filter(type != "Flom") %>%
  filter(type != "Flof")
mfattacktable <- table(mf$attack)
prop.table(mfattacktable)
chisq.test(mfattacktable, correct = T)
binom.test(mfattacktable, p = 0.5)

# Chi-squared approximation may be incorrect
# 	Chi-squared test for given probabilities
# 
# data:  mfattacktable
# X-squared = 2.7778, df = 1, p-value = 0.09558

fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(attack != 0)
fmlfattacktable <- table(fmlf$attack)
prop.table(fmlfattacktable)
chisq.test(fmlfattacktable)
binom.test(fmlfattacktable, p = 0.5)

# 	Chi-squared test for given probabilities
# 
# data:  fmlfattacktable
# X-squared = 13.714, df = 1, p-value = 0.0002128

fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(attack != 0) %>%
  filter(type != "Flom")
fmlfattacktable <- table(fmlf$attack)
prop.table(fmlfattacktable)
chisq.test(fmlfattacktable)
binom.test(fmlfattacktable)

# 	Chi-squared test for given probabilities
# 
# data:  fmlfattacktable
# X-squared = 10.889, df = 1, p-value = 0.0009674

fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(attack != 0) %>%
  filter(type != "Flom") %>%
  filter(type != "Flof")
fmlfattacktable <- table(fmlf$attack)
prop.table(fmlfattacktable)
chisq.test(fmlfattacktable)
binom.test(fmlfattacktable)


# 	Chi-squared test for given probabilities
# 
# data:  fmlfattacktable
# X-squared = 8.0667, df = 1, p-value = 0.004509

mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(attack != 0) 
mmlfattacktable <- table(mmlf$attack)
prop.table(mmlfattacktable)
chisq.test(mmlfattacktable)
binom.test(mmlfattacktable, p = 0.5)

# 	Chi-squared test for given probabilities
# 
# data:  mmlfattacktable
# X-squared = 0.037037, df = 1, p-value = 0.8474

mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(attack != 0) %>%
  filter(type != "Flom")
mmlfattacktable <- table(mmlf$attack)
prop.table(mmlfattacktable)
chisq.test(mmlfattacktable)
binom.test(mmlfattacktable, p = 0.5)

# Chi-squared approximation may be incorrect
# 	Chi-squared test for given probabilities
# 
# data:  mmlfattacktable
# X-squared = 0, df = 1, p-value = 1

mmlf <- testcategories_nona %>% 
  filter(test == "M/MLF") %>%
  filter(attack != 0) %>%
  filter(type != "Flom") %>%
  filter(type != "Flof")
mmlfattacktable <- table(mmlf$attack)
prop.table(mmlfattacktable)
chisq.test(mmlfattacktable)

# Chi-squared approximation may be incorrect
# 	Chi-squared test for given probabilities
# 
# data:  mmlfattacktable
# X-squared = 0.2, df = 1, p-value = 0.6547
```

What if we just look at species?
```{r}
mf <- testcategories_nona %>% 
  filter(test == "M/F") %>%
  filter(attack != 0) %>%
  mutate(species = if_else(type == "Ama", "other",
                                      if_else(type == "Other", "other", "florisuga")))

mfattacktable <- table(mf$species, mf$attack)
prop.table(mfattacktable, m=1)
fisher.test(mfattacktable)
prop.test(mfattacktable, correct = T)
# data:  mfattacktable
# X-squared = 9.6416e-31, df = 1, p-value = 1
(7+2)/(7+2+17+7)

fmlf <- testcategories_nona %>% 
  filter(test == "F/MLF") %>%
  filter(attack != 0) %>%
  mutate(species = if_else(type == "Ama", "other",
                                      if_else(type == "Other", "other", "florisuga")))

fmlfattacktable <- table(fmlf$species, fmlf$attack)
prop.table(fmlfattacktable, m=1)
fisher.test(fmlfattacktable)
prop.test(fmlfattacktable, correct = T)
# data:  fmlfattacktable
# X-squared = 0.31425, df = 1, p-value = 0.5751
(13+2)/(13+2+20+7)

mmlf <- testcategories_nona %>%
  filter(test == "M/MLF") %>%
  filter(attack != 0) %>%
  mutate(species = if_else(type == "Ama", "other",
                                      if_else(type == "Other", "other", "florisuga")))

mmlfattacktable <- table(mmlf$species, mmlf$attack)
prop.table(mmlfattacktable, m=1)
fisher.test(mmlfattacktable)
prop.test(mmlfattacktable, correct = T)
# data:  mmlfattacktable
# X-squared = 6.3196e-31, df = 1, p-value = 1
(2+3)/(2+3+11+11)

(2+3+13+2+7+2)/(2+3+11+11+13+2+20+7+7+2+17+7)
```

This looks at sex behavior of males toward the mounts. First need to shape the data a bit 
```{r}
#Find all the instances where courtship or copulation happened, and take out the ones where it didn't
courtscops_nona <- testcategories_nona %>%
  mutate(sumthing = if_else(court == 0 & cop == 0, 0, 1)) %>%
  filter(sumthing == 1) %>%
  select(date, trial, test, type, court, cop) %>%
  replace_na(list(court = 0, cop = 0))

#This combines courtship and copulation. If courtship happened, it takes the courtship choice. If not, but copulation happened, it takes the copulation choice
courtscops_combine <- courtscops_nona %>%
  mutate(combine = if_else(court != 0, court, cop))
```

Who did androchromes choose to direct sexual behavior toward? 
```{r}
#All chose F first in M/F tests
mf <- courtscops_combine %>% filter(test == "M/F") 
mfcourtcoptable <- table(mf$combine)
prop.table(mfcourtcoptable)
binom.test(mfcourtcoptable, p = 0.5)

mf_fortest <- c(12, 0)
binom.test(mf_fortest, p = .5)

# 	Chi-squared test for given probabilities
# 
# data:  mf_fortest
# X-squared = 12, df = 1, p-value = 0.000532

#All chose F first in M/MLF
fmlf <- courtscops_combine %>% filter(test == "F/MLF")

mmlf_fortest <- c(16, 0)
binom.test(mmlf_fortest, p = .5)

# 	Chi-squared test for given probabilities
# 
# data:  mmlf_fortest
# X-squared = 16, df = 1, p-value = 6.334e-05

#Some chose MLF and some chose M in M/MLF tests
mmlf <- courtscops_combine %>% 
  filter(test == "M/MLF")
mmlfcourtcoptable <- table(mmlf$combine)
prop.table(mmlfcourtcoptable)
chisq.test(mmlfcourtcoptable)
binom.test(mmlfcourtcoptable)

# Chi-squared approximation may be incorrect
# 	Chi-squared test for given probabilities
# 
# data:  mmlfcourtcoptable
# X-squared = 0.5, df = 1, p-value = 0.4795
```

##All interactions with Mounts

Shape the data and change Left and Right to the actual mount types
```{r}
#How many individual locations are in the dataset?
n_distinct(behaviors$Obs_id)
#26 different locations

#Add some useful columns
behaviorsp <- behaviors %>% 
  #Make a new column with the species but no number
  mutate(species = substr(subject, 1, 4)) %>%
  #Group together everything that's not a jacobin for the species
  mutate(species_cat = if_else(species == "FloM", "Male", 
                              if_else(species == "FloF", "Female", "Other"))) %>%
  #Group together everything that's not a jacobin for the modifier
  mutate(modifiers_cat = if_else(modifiers == "FloM", "Male", 
                                 if_else(modifiers == "FloF", "Female", "Other")))


#Make datasets of all the R and L mount types in each trial
sidesR <- trialsides_sides %>% select(Obs_id, Right)
sidesL <- trialsides_sides %>% select(Obs_id, Left)

#Seperate R and L interactions 
firstreactionR <- behaviorsp %>% filter(modifiers == "Right")
firstreactionL <- behaviorsp %>% filter(modifiers == "Left")

#Join the mount type with all the R interactions for each trial
firstreactionR <- firstreactionR %>%
  inner_join(sidesR, by = "Obs_id") %>%
  select(-modifiers) %>%
  rename(mount = Right)

#Join the mount type with all the L interactions for each trial
firstreactionL <- firstreactionL %>%
  inner_join(sidesL, by = "Obs_id") %>%
  select(-modifiers) %>%
  rename(mount = Left)

#Bind the R and L interactions together again for a complete dataset
firstreactionmounts <- rbind(firstreactionL, firstreactionR)
```

Who did males tend to chase and who chased them? 
```{r}
#First, how many chases overall involved an androchrome? 
behaviorsp %>% 
  arrange(start)

behaviorsp %>% 
  filter(behavior == "Chase") %>%
  filter(species == c("FloM", "Flof"))

#Filter all the interactions where a male jacobin was the chaser
mchaser <- behaviorsp %>% filter(species == "FloM") %>% filter(behavior == "Chase")
#228 instances of a male chasing another bird
mchaser %>% filter(modifiers == "FloM")
#114 of those are against another Flom
mchaser %>% filter(modifiers == "FloF")
#72 of those are against another Flof
mchaser %>% filter(modifiers != "FloM") %>% filter(modifiers != "FloF")
#39 of those are against non-jacobins or unknowns

mchasee <- behaviorsp %>%
  filter(species == "FloM") %>%
  filter(behavior == "Chased by")
#157 instances of a male being chased by another bird
mchasee %>% filter(modifiers == "FloM")
#116 of those are from other Floms
mchasee %>% filter(modifiers == "FloF")
#2 of those are from Flofs
mchasee %>% filter(modifiers != "FloM") %>% filter(modifiers != "FloF")
#39 of those are from something else or unknown

#chaser to chased ratio of heterochromes
nrow(mchaser)/nrow(mchasee)
```

Who did females tend to chase and who chased them? 
```{r}
fchaser <- behaviorsp %>% filter(species == "FloF") %>% filter(behavior == "Chase")
#7 instances of a female chasing another bird
# 4 of those chases were towards Floms, 2 towards Flofs, and 1 toward Ama

fchasee <- behaviorsp %>%
  filter(species == "FloF") %>%
  filter(behavior == "Chased by")
#87 instances of a female being chased by another bird
fchasee %>% filter(modifiers == "FloM")
#73 of those are from Floms
fchasee %>% filter(modifiers == "FloF")
#3 of those are from Flofs
fchasee %>% filter(modifiers != "FloM") %>% filter(modifiers != "FloF")
#10 of those are from something else or unknown

#chaser to chased ratio of heterochromes
nrow(fchaser)/nrow(fchasee)
```

Plots
```{r}
chases <- behaviorsp %>%
  filter(behavior %in% c("Chase", "Chased by")) %>%
  filter(!is.na(modifiers_cat)) %>%
  filter(!is.na(species_cat))

ggplot(chases, aes(behavior, fill = modifiers_cat)) +
  geom_bar(position = "stack", width = 0.7) +
  #scale_x_discrete(limits = c("Male", "Female", "Other")) +
  facet_wrap(~ factor(species_cat, levels = c("Male", "Female", "Other"))) +
  theme_minimal()

ggplot(chases, aes(behavior, fill = modifiers)) +
  geom_bar(position = "stack", width = 0.7) +
  #scale_x_discrete(limits = c("Male", "Female", "Other")) +
  facet_wrap(~ factor(species)) +
  theme_minimal()

chases %>% filter(behavior == "Chased by") %>% filter(modifiers_cat == "Other") %>% filter(species_cat == "Male") %>% arrange(Obs_date)
chases %>% filter(behavior == "Chase") %>% filter(modifiers_cat == "Other") %>% filter(species_cat == "Male") %>% arrange(Obs_date)

ggplot(chases, aes(species_cat, fill = behavior)) +
  geom_bar(position = "dodge", width = 0.5) +
  scale_fill_manual(values=c("#00CC66","#0066CC"), labels = c("Chaser", "Chasee"))+
  scale_x_discrete(limits = c("Male", "Female")) +
  labs(x = " ", y = "Instances") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 14)) +
  theme(legend.title = element_blank(), legend.text = element_text(size = 16), legend.position = c(0.8, 0.8)) 

```

A sanity check. I was concerned why it looked like FloM chases of AmaT and AmaT chases of FloM were nearly equal. So I looked at the data closer
```{r}
#Instances where FloM chased AmaT
chases %>% filter(species == "FloM") %>% filter(behavior == "Chase") %>% filter(modifiers == "AmaT")
#Instances where AmaT chased FloM
chases %>% filter(species == "FloM") %>% filter(behavior == "Chased by") %>% filter(modifiers == "AmaT")

#If somehow there was a mistake where these two were linked, then the Chase and Chase by incidences should be at the same time. They are not, which means the similar number of chases is a coincidence.  
amat %>% filter(behavior %in% c("Chase", "Chased by"), modifiers == "AmaT") %>% arrange(start, modifiers)
```


```{r}
fmounts <- firstreactionmounts %>% filter(mount == "F")
#2239 total interactions with an F mount

fsexual <- fmounts %>%
  filter(behavior %in% c("U Dive" ,"Copulation"))
#28 of those instances are sexual (u dives or copulation)

faggression <- fmounts %>%
  filter(behavior %in% c("Peck", 
                         "Slam", 
                         "Mount Attack"#, 
                         #"Pass", 
                         #"Lunge"
                         )) %>%
  filter(species_cat == "Male")
#510 of the instances are aggressive in nature (pecks, slams, lunges, passes, and mount attacks)

f_sex_vs_aggression <- rbind(fsexual, faggression)

sex_vs_aggression <- f_sex_vs_aggression %>%
  mutate(interact = if_else(behavior %in% c("Peck", "Slam", "Mount Attack"), "aggression", "courtship"))

g <- ggplot(f_sex_vs_aggression, aes(interact)) +
  geom_bar(width = 0.3, fill = "#3399FF") +
  labs(x = " ", y = "Instances") +
  #theme_minimal() +
  theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 14)) +
  theme(legend.title = element_blank(), legend.text = element_text(size = 16), legend.position = c(0.8, 0.8)) 

g

```

```{r}


mmounts <- firstreactionmounts %>% filter(mount == "M")
#2687 total interactions with an M mount

msexual <- mmounts %>%
  filter(behavior %in% c("U Dive" ,"Copulation"))
#80 of those instances are sexual (u dives or copulation)

maggression <- mmounts %>%
  filter(behavior %in% c("Peck", "Slam", "Mount Attack"))
#943 of the instances are aggressive in nature (pecks, slams, lunges, passes, and mount attacks)

mlfmounts <- firstreactionmounts %>% filter(mount == "MLF")
#2212 total interactions with an MLF mount

mlfsexual <- mlfmounts %>%
  filter(behavior %in% c("U Dive" ,"Copulation"))
#28 of those instances are sexual (u dives or copulation)

mlfaggression <- mlfmounts %>%
  filter(behavior %in% c("Peck", "Slam", "Mount Attack"))
#731 of the instances are aggressive in nature (pecks, slams, lunges, passes, and mount attacks)
```

